---
title: "creating-expert-demos"
author: "Grace Douglas"
date: "2024-04-03"
output: html_document
---

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(naniar)
library(readr)
library(stars)
library(spacetime)
library(magrittr)
library(sftime)
library(stplanr)
library(lubridate)
library(dplyr)
library(xts) # For handling time series data
library(lubridate) # For date-time manipulation
knitr::opts_chunk$set(out.width="800px", dpi=80)
library(tidyr)

setwd("~/imitation/mal-proj/mal/")
bouts <- st_read("~/imitation/mal-proj/april 11/bouts-1ft-samplingv2.shp")  %>%
  st_transform(crs = 4326)

```



```{r}
# Rename duplicated column names to make them unique
bouts = bouts %>%
  rename(ft_s = tf_s)

#what frequency are we at
sampling_rates <- bouts %>%
  group_by(rowid) %>%
  summarise(sampling_rate = mean(cngfeet / ft_s, na.rm = TRUE))

bouts = bouts %>% filter(!is.na(id))
bouts


# Convert 'tmin' and 'tmax' to POSIXct format
bouts$tmin <- ymd_hms(bouts$tmin)
bouts$tmax <- ymd_hms(bouts$tmax)

# Calculate the time difference between 'tmin' and 'tmax'
bouts$duration_sec <- as.numeric(difftime(bouts$tmax, bouts$tmin, units = "secs"))

bouts = bouts[1:200,]

# Create a sequence of timestamps for each 'id' based on 'tmin', 'tmax', and 'fid'
bouts <- bouts %>%
  group_by(id) %>%
  mutate(
    timestamp = purrr::
      map2(tmin, tmax, ~seq.POSIXt(from = .x, to = .y, length.out = length(fid))),
    timestamp = unlist(timestamp)
  ) %>%
  ungroup()

# Calculate the time difference between consecutive timestamps
bouts <- bouts %>%
  arrange(id, fid) %>%
  mutate(
    time_diff_sec = c(0, diff(timestamp)),
    speed = mean_speed * 1000 / 3600  # Convert mean speed from kilometers per hour to meters per second
  )

# Interpolate the distance covered for each 'fid'
bouts$distance_covered <- ave(bouts$cngfeet, bouts$id, FUN = cumsum)

# Calculate the distance covered in each time interval
bouts$distance_diff <- c(0, diff(bouts$distance_covered))

# Calculate the time taken to cover each distance increment
bouts$time_increment <- bouts$distance_diff / bouts$speed

# Calculate the cumulative time taken
bouts$cumulative_time <- cumsum(bouts$time_increment)

# Add the cumulative time to the initial timestamp to get the extrapolated timestamp for each 'fid'
bouts$extrapolated_timestamp <- bouts$timestamp[1] + bouts$cumulative_time

# Select relevant columns
result <- bouts %>%
  select(id, fid, extrapolated_timestamp, speed) %>%
  distinct()  # Remove duplicated rows

# View the resulting dataframe
print(result)

```


## reconstruct each bout using tmin, tmax pairs, duration_m, n and sf.geo

[Linestring to points](https://docs.ropensci.org/stplanr/reference/line2points.html) 
The points corresponding with a given line, i, will be $(2*i):((2*i)+1)$

order of operations:
1. make loess curve for each linestring
2. resample at 1 second intervals from lowess curve for each linestring
3. compute speed between 1 second lowess curve points (average speed will be same as original traversal, but speed along points will vary)

Loess stands for locally estimated scatterplot smoothing (lowess stands for locally weighted scatterplot smoothing) and is one of many non-parametric regression techniques, but arguably the most flexible.

```{r, fig.width=10, fig.height=10}
bout = walk
bout_pts <- line2pointsn(walk) #row.names are pts in linestring
plot(bout$geometry)
plot(bout_pts$geometry, add = TRUE)

time_step <- 1 # in seconds
#walk$datetime <- ymd_hms(walk$tmin)

```


## snap gps traces to osm grid - DO NOT DO THIS.

> snapping to osm conforms to osm grid --> loosing all individual preferential micro route-planning (i.e., jaywalk)


```{r, eval = FALSE}
library(sf)
library(dplyr)
library(tidygeocoder)
library(mapboxapi)
library(hereR)
library(osrm)
library(leaflet)
token = "pk.eyJ1IjoiZ3JhY2tpdGgiLCJhIjoiY2x1cm56bDZoMDljNTJtbDYzZGk1NGdmMCJ9.dYt96hiPMofNjlPoF5VAYA"
#mb_access_token(token, install = TRUE, overwrite=TRUE)
#Sys.getenv("MAPBOX_PUBLIC_TOKEN")
get_mb_access_token()

osroute <- osrm::osrmRoute(loc = walk,
                           returnclass = "sf")

#must split into rows before feeding
mroute <- mapboxapi::mb_directions(input_data = walk,
                                   profile = "walking")

leaflet(osroute) %>%
  addMapboxTiles(
    style_id = "light-v9",
    username = "mapbox"
  ) %>%
  addPolylines()
```



## interpolate points to get a timeseries by time-step

```{r}
interpolated_points <- st_interpolate(walk$geometry, n = 100)

```



```{r}
glimpse(traj_raw_slow)
lo <- loess(y~duration_min)
plot(x,y)
lines(predict(lo), col='red', lwd=2)
```

