---
title: "visualizing-expert-trajectories-action"
author: "Grace Douglas"
date: "2024-04-03"
output: html_document
---

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(naniar)
library(readr)
library(stars)
library(spacetime)
library(magrittr)
library(sftime)
library(stplanr)
library(lubridate)

library(xts) # For handling time series data
library(lubridate) # For date-time manipulation
knitr::opts_chunk$set(out.width="800px", dpi=80)

setwd("~/imitation/mal-proj/mal/")
```

## bring trajectories in and combine

```{r}
# ids to keep from above cleaning

action = st_read("~/imitation/mal-proj/greenlake/action-gl.shp") %>%
  st_transform(int, crs = 4326) %>% 
  tibble::rowid_to_column()
names(action) = tolower(names(action))

trac <- st_read("~/imitation/mal-proj/greenlake/trac-gl.shp") %>%
  st_transform(crs = 4326) %>%
  tibble::rowid_to_column() 
names(trac) = tolower(names(trac))

action = action %>%
  dplyr::select(-run, -axis1_mean, -median_spe, -is_tour, -n_records, -pct_record, -dist_m_sta, -n_records_) %>%
  rename(boutnum__1 =lmpaboutnu)

trac = trac %>%
  dplyr::select(-has_tlog, -overlaps_w, -overlaps_n, -overlaps_a, -is_complet, -has_comple, -valid_10h_, -cpm_mean)%>%
  rename(tmin = time_pa_10,
         tmax = time_pa_01,
         calendar_d = jdaya)

# Deactivate geometry columns
action_df <- action %>% st_drop_geometry()
trac_df <- trac %>% st_drop_geometry()

#trac_df_reordered <- trac_df[, action_df$columns]
# walk <- merge(action_df, trac_df)
# walk <- sp::merge(action_df, trac_df)

walk = rbind(action, trac)

# Reactivate geometry columns
st_geometry(walk) <- c(action$geometry, trac$geometry)

```


## fix missing geometry
```{r}
walk = walk %>%
  st_transform(int, crs = 4326)
#dim : 42699    14

#filter empty geometries out by surrogate variable radius_mbc as to not disrupt current geometry
walk = walk %>% filter(!is.na(radius_mbc))
```

## filter transit users out
```{r}
walk = walk %>% filter(is_transit != "1" & is_transit != "T") %>% dplyr::select(-is_transit)

dim(walk)
length(unique(walk$id))
```


# we must select a number of experts
## but what is the task? 

- discriminator can distinguish between expert demonstration and 
- preferential path-planning: 
  - runners vs walkers? 
  - pt.a -> pt.b: go *around* greenlake
  - classify: `util` variable 
5 levels: 
  nonwalk
  rec_gps_loop
  rec_tour      
  unknown    
  util

```{r}
# walk %>%
#   group_by(id) %>%
#   summarise(n = n()) %>%
#   arrange(desc(n), .by_group = TRUE) %>%
#   ggplot(mapping = aes(x = reorder(factor(id), -n), y = n)) + 
#   geom_col() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))

walk <- st_zm(walk, drop = TRUE) # Drop Z coordinate
#st_write(walk, "~/imitation/mal-proj/mal/for-smoothing-qgis.shp")
```


# there is not much speed information missing, but we should impute the 0 values
```{r}
library(mice)
walk$mean_speed[walk$mean_speed==0] = NA # specify mean_speed == 0 as NA
walk$walk_type = as.factor(walk$walk_type) # factorize
walk$rec_util = as.factor(walk$rec_util) # factorize
walk$id = as.factor(walk$id)
walk$rowid = as.factor(walk$rowid)
walk_mice = walk %>% dplyr::select(-id) %>% st_drop_geometry()


base <- mice(walk_mice[,-1], maxit=0) # run once with 0 imputations to compute the predM dichotomous matrix
predM = base$predictorMatrix
meth = base$method
imp = mice(walk_mice[,-1], maxit = 5, 
             predictorMatrix = predM, 
             method = meth, print =  TRUE)
walk_impute = mice::complete(imp, action = 5L, include = FALSE)
dim(walk_impute)
dim(walk)

walk_impute$rowid = walk$rowid
walk_impute$id = walk$id
st_geometry(walk_impute) = st_geometry(walk)

#st_write(walk_impute, "~/imitation/mal-proj/mal/fromR-for-qgis-smoothing-v2.shp")

```




